Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)

### 2014-04-24 WSSSPE1_report/WSSSPE1_summary_report.tex
some fixes

% Why are we classifying software like above?
This classification is motivated by our argument that the two kinds of
software projects wildly vary in factors such as scope, purpose and
usage. The development and management of each kind is significantly
different. Consequently, the sustainability challenges faced by them
differ and must be treated separately. For instance, the challenges
faced by a gateway software development project such as
\toolname{CIPRES} or visualization software pinroducts such as
\toolname{VisIT} or \toolname{ParaView} are distinct to a niche software for
\textit{ab initio} modeling and simulation such as \toolname{VASP} or
\toolname{Quantum Espresso}. \note{cite SW}

@@

On the other hand, utility software presents opportunities to be
usable by a larger community making its undertaking and development an
attractive pursuit.  It is generally more visible in community which
in turn leads to a broader and deeper participation. For instance,
promoting collaborations across the breadth (e.g., different
departments) and depth (e.g., stakeholders within a department) of
community, one of the key ingredients of a sustainable
process. Successful utility projects reap high rewards and have a
longer usage span. Development process becomes user-driven and
self-sustaining.

@@

Finally, it appears that the career options of developers from
two kinds of software projects differ significantly. In order to
further investigate the much discussed issue of career paths for
developers of these software, we conducted a small survey of the cases
discussed above with respect to the current status of its core
developers. We looked at the workshop submissions and the original
cited papers for the respective software. Where available, we looked
at the web pages and source code of the earliest versions to find the
original developers. We included projects older than 5 years and
excluded relatively new projects such as \toolname{SciPy} and
\toolname{numFOCUS} from this survey. We find that in a majority of
scientific software, more than one original developer are still
around, in some cases after more than 10 years (e.g.,
\toolname{R/qtl}).  Whereas, in the case of utility software projects,
it appears that at least one original developer is still around, often
in a part-time or ``benefactor'' role. In other cases, it appears that
the original developers have went on to assume a larger role with the
same project.

>>>>

% Why are we classifying software like above?
This classification is motivated by our argument that the two kinds of
software projects wildly vary in factors such as scope, purpose and
usage. The development and management of each kind is significantly
different. Consequently, the sustainability challenges faced by them
differ and must be treated separately. For instance, the challenges
faced by a gateway software development project such as
\toolname{CIPRES} or visualization software products such as
\toolname{VisIT} or \toolname{ParaView} are distinct to a niche software for
\textit{ab initio} modeling and simulation such as \toolname{VASP} or
\toolname{Quantum Espresso}. \note{cite SW}

@@

On the other hand, utility software presents opportunities to be usable by a
larger community making its undertaking and development an attractive pursuit.
It is generally more visible in community which in turn leads to a broader and
deeper participation. For instance, it helps promoting collaborations across
the breadth (e.g., different departments) and depth (e.g., stakeholders within
a department) of community, one of the key ingredients of a sustainable
process. Successful utility projects reap high rewards and have a longer usage
span. Development process becomes user-driven and self-sustaining.

@@

Finally, it appears that the career options of developers from
two kinds of software projects differ significantly. In order to
further investigate the much discussed issue of career paths for
developers of these software, we conducted a small survey of the cases
discussed above with respect to the current status of its core developers. We
looked at the workshop submissions and the original cited papers for the
respective software. Where available, we looked at the web pages and source
code of the earliest versions to find the original developers. We included
projects older than 5 years and excluded relatively new projects such as
\toolname{SciPy} and \toolname{numFOCUS} from this survey. We find that in a
majority of scientific software, more than one original developer are still
around, in some cases after more than 10 years (e.g., \toolname{R/qtl}).
Whereas, in the case of utility software projects, it appears that at least one
original developer is still around, often in a part-time or ``benefactor''
role. In other cases, it appears that the original developers have went on to
assume a larger role (eg. manager) within the same project.
